---
title: "eeg-sim"
output: html_document
date: "2022-10-31"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Here is code to stick in an R notebook:



---
title: "Cluster onsets: simulated signals"
author: "Guillaume A. Rousselet"
date: "`r Sys.Date()`"
output:
  pdf_document:
    fig_caption: no
    number_sections: no
    toc: yes
    toc_depth: 2
    # github_document:
    # html_preview: yes
    # toc: yes
    # toc_depth: 2
---



# Dependencies
```{r}
library(ggplot2)
library(tibble)
library(tidyverse)
#library(primer) # for one_over_f function
# library(permuco)
```



Edit `one_over_f` function to control variance.
Original function is available on [GitHub](https://github.com/HankStevens/primer).
Copyright Hank Stevens.



```{r}
one_over_f <- function(gamma = 1, N = 200, outvar = 1){
    N.2 <- N/2
    sine.waves <- matrix(NA, nrow = N, ncol = N.2)
    steps = 2 * pi * (1:N)/N
    phase <- stats::runif(N.2, 0, 2 * pi)
    for (i in 1:N.2) {
        freq <- i
        weight <- 1/(freq^gamma)
        y <- weight * sin(freq * steps + phase[i])
        sine.waves[, i] <- y
    }
    out <- rowSums(sine.waves)
    # force var = outvar and mean = 0
    out <- out - mean(out);
    out <- out / sd(out);
    out <- out * sqrt(outvar);
    return(out)
}
```




# Make template



```{r}
# true onset = 160 ms, F=17, max at F=26
true_onset <- 160
Xf <- seq(0, 500, 10)
Nf <- length(Xf)
temp1 <- vector(mode = "numeric", length = Nf)
erp <- dnorm(seq(-1.5, 1.5, length.out=21), 0, 1)
erp <- erp - min(erp)
erp <- erp / max(erp)
temp2 <- c(rep(0, 15), erp, rep(0, 15))
```



## Illustrate template
```{r}
df <- tibble(x = Xf,
             y = temp2)
ggplot(df, aes(x, y)) +
  geom_line()
```



# Generate noise



## For one gamma value
```{r}
Nt <- 5 # number of trials
gsp <- 0.7 # gamma spectral power
resmat <- matrix(data = 0, nrow = Nt, ncol = Nf)
for(T in 1:Nt){
  resmat[T,] <- one_over_f(gamma = gsp, Nf)
}



df <- tibble(x = rep(Xf, each = Nt),
             y = as.vector(resmat),
             trials = factor(rep(seq(1,Nt), Nf)))
ggplot(df, aes(x, y, group = trials, colour = trials)) +
  geom_line()
```



## For multiple gamma values
```{r}
gsp <- seq(0, 1, 0.2) # gamma spectral power
outvar <- 1 # noise variance
Nt <- length(gsp)
resmat <- matrix(data = 0, nrow = Nt, ncol = Nf)
for(T in 1:Nt){
  set.seed(1) # same random seed to better compare trials
  resmat[T,] <- one_over_f(gamma = gsp[T], Nf, outvar = outvar)
}



df <- tibble(x = rep(Xf, each = Nt),
             y = as.vector(resmat),
             gamma = as.factor(rep(gsp, Nf)))


ggplot(df, aes(x, y, group = gamma)) +
  geom_line(size = 1) +
  # scale_colour_viridis_d(guide = guide_coloursteps(barwidth = 20,
  #                                                  barheight = .5,
  #                                                  show.limits = T),
  #                        # breaks = seq(0, 1, 0.1),
  #                        # labels = c("", 0.2, "", 0.4, "", 0.6, "", 0.8, "", 1.0)
  #                        )+
  theme_minimal()+
  theme(legend.position = "bottom")+
  facet_wrap(vars(gamma))
```
## Illustrate signals + noise



```{r}
# set.seed(666)
gsp <- 0.7 # gamma spectral power
outvar <- 0.1 # noise variance
cond2 <- temp2 + one_over_f(gamma = gsp, Nf, outvar = outvar)
cond1 <- temp1 + one_over_f(gamma = gsp, Nf, outvar = outvar)



df <- tibble(x = rep(Xf,2),
             y = c(cond1, cond2),
             conds = factor(rep(c("Condition 1", "Condition 2"), each = Nf)))
ggplot(df, aes(x, y, colour = conds)) +
  geom_line()



# Nt <- 12
# resmat.sig <- matrix(data = 0, nrow = Nt, ncol = Nf)
# resmat.noi <- matrix(data = 0, nrow = Nt, ncol = Nf)
# for(T in 1:Nt){
#   resmat.sig[T,] <- temp2 + one_over_f(gamma = gsp, Nf, outvar = outvar)
#   resmat.noi[T,] <- temp1 + one_over_f(gamma = gsp, Nf, outvar = outvar)
# }
#
# df <- tibble(x = rep(rep(Xf, each = Nt),2),
#              y = c(as.vector(resmat.sig), as.vector(resmat.noi)),
#              trials = factor(rep(1:Nt, Nf*2)),
#              cat = factor(rep(c("Signal", "Noise"), each = Nt*Nf)))
# ggplot(df, aes(x, y, group = trials, colour = cat)) +
#   geom_line(size = 1) +
#   facet_wrap(vars(trials))
```



## Illustrate averages



```{r}
# set.seed(666)
Nt <- 50
Nts <- c(10,25,50,100,250,500) # number of trials
gsp <- 0.7 # gamma spectral power
outvar <- 1 # noise variance

dat = vector("list", length = length(Nt))

for (i in 1:length(Nts)) {

cond1 <- matrix(0, nrow = Nts[i], ncol = Nf)
cond2 <- matrix(0, nrow = Nts[i], ncol = Nf)

  for(t in 1:Nts[i]){
    
  cond2[t,] <- temp2 + one_over_f(gamma = gsp, Nf, outvar = outvar)
  cond1[t,] <- temp1 + one_over_f(gamma = gsp, Nf, outvar = outvar)  
  
  }

cond1_mn <- apply(cond1, 2, mean)
cond2_mn <- apply(cond2, 2, mean)

dat[[i]] <- tibble(x = rep(Xf,2),
             y = c(cond1_mn, cond2_mn),
             conds = factor(rep(c("Condition 1", "Condition 2"), each = Nf)), 
             n_trial = Nts[i])

}
```


```{r}
  
eeg_df <- dat |> 
  bind_rows()
  

eeg_df |> 
  ggplot(aes(x, y, colour = conds)) +
  geom_line() +
  geom_vline(xintercept = true_onset)+
  facet_wrap(~n_trial, scales =  "free")+
  theme_minimal()
```




```{r}
Nt <- 50
Nts <- c(10,25,50,100,250,500) # number of trials
gsp <- 0.7 # gamma spectral power
outvar <- 1 # noise variance

alpha = 0.05

t_dat <- vector("list", length = length(Nts))

for (i in 1:length(Nts)){

  p_values <- array(NA, Nf)
  mn_values <- array(NA, Nf) # mean difference values
  
  cond1 <- matrix(0, nrow = Nts[i], ncol = Nf)
  cond2 <- matrix(0, nrow = Nts[i], ncol = Nf)

  for(t in 1:Nts[i]){
      
    cond2[t,] <- temp2 + one_over_f(gamma = gsp, Nf, outvar = outvar)
    cond1[t,] <- temp1 + one_over_f(gamma = gsp, Nf, outvar = outvar)  
  }
  
  for(x in 1:(Nf)){
    t_test_values <- t.test(cond2[,x], cond1[,x], var.equal = FALSE)
    p_values[x] <- t_test_values[["p.value"]]
    mn_values[x] <- abs(t_test_values[["estimate"]][[1]] - t_test_values[["estimate"]][[2]])
  }

t_dat[[i]] <- tibble(
  p = p_values,
  time = Xf,
  mn_diff = mn_values,
  n_trials = Nts[i]
  ) |>
  mutate(sig = if_else(p < alpha/Nts[i], 0, NULL))

}

```


```{r}
t_df <- t_dat |> 
  bind_rows()
```


```{r}
rope_noise <- quantile(mn_values[1:16], probs = c(0.025, 0.975)) # from 10-160 milliseconds - pre-stimulus
```


Plotting P-values
```{r}
t_df |> 
  ggplot(aes(x = time, y = -log(p)))+
  geom_point(aes(colour = p < (alpha/ Nf)), show.legend = F, size = 2)+
  geom_line() +
  geom_hline(aes(yintercept = -log(alpha/ Nf)))+
  theme_minimal()+
  facet_wrap(~n_trials, scales = "free_y")


t_df |> 
  ggplot(aes(x = time, y = p))+
  geom_line()+
  geom_point(aes(colour = p < (alpha/ Nf)), show.legend = F, size = 2)+
  geom_hline(aes(yintercept = alpha/ Nf))+
  theme_minimal()+
  facet_wrap(~n_trials, scales = "free_y")
```

```{r}
t_df |> 
  select(-sig) |> 
  pivot_wider(names_from = time, values_from = c(p, mn_diff)) |> 
  select(contains("p_")) |> 
  rename_with(.cols = everything(), .fn =  ~ str_c(str_extract_all(.x, pattern = "\\d+"))) |> 
  corrr::correlate(use = "pairwise.complete.obs", method = "pearson") |> 
  corrr::autoplot(triangular = "full", method = "Identity")+
  scale_x_discrete(labels = str_c(seq(0, 500, 50)), breaks = seq(0, 500, 50))+
  scale_y_discrete(labels = str_c(seq(0, 500, 50)), breaks = seq(0, 500, 50))
  
t_df |> 
  select(-sig) |> 
  pivot_wider(names_from = time, values_from = c(p, mn_diff)) |> 
  select(contains("mn_")) |> 
  rename_with(.cols = everything(), .fn =  ~ str_c(str_extract_all(.x, pattern = "\\d+"))) |> 
  corrr::correlate(use = "pairwise.complete.obs", method = "pearson") |> 
  corrr::autoplot(triangular = "full", method = "Identity")+
  scale_x_discrete(labels = str_c(seq(0, 500, 50)), breaks = seq(0, 500, 50))+
  scale_y_discrete(labels = str_c(seq(0, 500, 50)), breaks = seq(0, 500, 50))
```


Plotting Mean Difference
```{r}
t_df |> 
  ggplot(aes(x = time, y = mn_diff))+
  geom_line()+
  geom_point(size = 2, show.legend = F, colour = "darkcyan")+
  theme_minimal()+
  # annotate("rect", 
  #          xmin = 0, 
  #          xmax = max(t_test_df$time), 
  #          ymin = rope_noise[1], 
  #          ymax = rope_noise[2], 
  #          alpha = .2)+
  geom_line(aes(x = time, y = sig), size = 1.5)+
  facet_wrap(~n_trials)


```


```{r}
cond1_df <- cond1[1:100, ] |> 
  as_tibble(.name_repair = ~ paste0(Xf))|> 
  pivot_longer(cols = everything(), names_to = "time", values_to = "milliv") |> 
  mutate(trial = rep(1:100, times = Nf))


cond2_df <- cond2[1:100, ] |> 
  as_tibble(.name_repair = ~ paste0(Xf))|> 
  pivot_longer(cols = everything(), names_to = "time", values_to = "milliv") |>   
  mutate(trial = rep(1:100, times = Nf))

cond_df <- bind_rows(cond1_df, cond2_df, .id = "cond") |> 
  group_by(time, cond) |> 
  mutate(mn = mean(milliv))
```


```{r}
cond_df |> 
  ggplot(aes(as.numeric(time), milliv, colour = as.factor(trial), group = trial))+
  geom_line(size = .4, alpha = .4, show.legend = F)+
  geom_line(aes(y = mn), size = 1, alpha = 1, colour = "slategray4")+
  facet_grid(rows = vars(cond))+
  scale_x_continuous("Time (ms)")+
  theme_minimal()
```


