---
title: "eeg-sim"
output: html_document
date: "2022-10-31"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Here is code to stick in an R notebook:



---
title: "Cluster onsets: simulated signals"
author: "Guillaume A. Rousselet"
date: "`r Sys.Date()`"
output:
  pdf_document:
    fig_caption: no
    number_sections: no
    toc: yes
    toc_depth: 2
    # github_document:
    # html_preview: yes
    # toc: yes
    # toc_depth: 2
---



# Dependencies
```{r}
library(ggplot2)
library(tibble)
#library(primer) # for one_over_f function
# library(permuco)
```



Edit `one_over_f` function to control variance.
Original function is available on [GitHub](https://github.com/HankStevens/primer).
Copyright Hank Stevens.



```{r}
one_over_f <- function(gamma = 1, N = 200, outvar = 1){
    N.2 <- N/2
    sine.waves <- matrix(NA, nrow = N, ncol = N.2)
    steps = 2 * pi * (1:N)/N
    phase <- stats::runif(N.2, 0, 2 * pi)
    for (i in 1:N.2) {
        freq <- i
        weight <- 1/(freq^gamma)
        y <- weight * sin(freq * steps + phase[i])
        sine.waves[, i] <- y
    }
    out <- rowSums(sine.waves)
    # force var = outvar and mean = 0
    out <- out - mean(out);
    out <- out / sd(out);
    out <- out * sqrt(outvar);
    return(out)
}
```




# Make template



```{r}
# true onset = 160 ms, F=17, max at F=26
true_onset <- 160
Xf <- seq(0, 500, 10)
Nf <- length(Xf)
temp1 <- vector(mode = "numeric", length = Nf)
erp <- dnorm(seq(-1.5, 1.5, length.out=21), 0, 1)
erp <- erp - min(erp)
erp <- erp / max(erp)
temp2 <- c(rep(0, 15), erp, rep(0, 15))
```



## Illustrate template
```{r}
df <- tibble(x = Xf,
             y = temp2)
ggplot(df, aes(x, y)) +
  geom_line()
```



# Generate noise



## For one gamma value
```{r}
Nt <- 5 # number of trials
gsp <- 0.7 # gamma spectral power
resmat <- matrix(data = 0, nrow = Nt, ncol = Nf)
for(T in 1:Nt){
  resmat[T,] <- one_over_f(gamma = gsp, Nf)
}



df <- tibble(x = rep(Xf, each = Nt),
             y = as.vector(resmat),
             trials = factor(rep(seq(1,Nt), Nf)))
ggplot(df, aes(x, y, group = trials, colour = trials)) +
  geom_line()
```



## For multiple gamma values
```{r}
gsp <- seq(0, 1, 0.2) # gamma spectral power
outvar <- 1 # noise variance
Nt <- length(gsp)
resmat <- matrix(data = 0, nrow = Nt, ncol = Nf)
for(T in 1:Nt){
  set.seed(1) # same random seed to better compare trials
  resmat[T,] <- one_over_f(gamma = gsp[T], Nf, outvar = outvar)
}



df <- tibble(x = rep(Xf, each = Nt),
             y = as.vector(resmat),
             gamma = factor(rep(gsp, Nf)))
ggplot(df, aes(x, y, group = gamma)) +
  geom_line(size = 1) +
  # scale_colour_viridis_d(end = 0.95)
  facet_wrap(vars(gamma))
```
## Illustrate signals + noise



```{r}
# set.seed(666)
gsp <- 0.7 # gamma spectral power
outvar <- 0.1 # noise variance
cond2 <- temp2 + one_over_f(gamma = gsp, Nf, outvar = outvar)
cond1 <- temp1 + one_over_f(gamma = gsp, Nf, outvar = outvar)



df <- tibble(x = rep(Xf,2),
             y = c(cond1, cond2),
             conds = factor(rep(c("Condition 1", "Condition 2"), each = Nf)))
ggplot(df, aes(x, y, colour = conds)) +
  geom_line()



# Nt <- 12
# resmat.sig <- matrix(data = 0, nrow = Nt, ncol = Nf)
# resmat.noi <- matrix(data = 0, nrow = Nt, ncol = Nf)
# for(T in 1:Nt){
#   resmat.sig[T,] <- temp2 + one_over_f(gamma = gsp, Nf, outvar = outvar)
#   resmat.noi[T,] <- temp1 + one_over_f(gamma = gsp, Nf, outvar = outvar)
# }
#
# df <- tibble(x = rep(rep(Xf, each = Nt),2),
#              y = c(as.vector(resmat.sig), as.vector(resmat.noi)),
#              trials = factor(rep(1:Nt, Nf*2)),
#              cat = factor(rep(c("Signal", "Noise"), each = Nt*Nf)))
# ggplot(df, aes(x, y, group = trials, colour = cat)) +
#   geom_line(size = 1) +
#   facet_wrap(vars(trials))
```



## Illustrate averages



```{r}
# set.seed(666)
Nt <- 50 # number of trials
gsp <- 0.7 # gamma spectral power
outvar <- 1 # noise variance
cond1 <- matrix(0, nrow = Nt, ncol = Nf)
cond2 <- matrix(0, nrow = Nt, ncol = Nf)
for(T in 1:Nt){
cond2[T,] <- temp2 + one_over_f(gamma = gsp, Nf, outvar = outvar)
cond1[T,] <- temp1 + one_over_f(gamma = gsp, Nf, outvar = outvar)  
}
cond1 <- apply(cond1, 2, mean)
cond2 <- apply(cond2, 2, mean)



df <- tibble(x = rep(Xf,2),
             y = c(cond1, cond2),
             conds = factor(rep(c("Condition 1", "Condition 2"), each = Nf)))
ggplot(df, aes(x, y, colour = conds)) +
  geom_line() +
  geom_vline(xintercept = true_onset)
```