---
title: "Power"
output: html_document
date: "2022-09-29"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(pwr)

```

Setting default values for the power calculation
```{r}
sample = 5000
sample_2 = 1000 # pwr calc doesn't like alpha being 0, so we stop earlier
alpha = 0.05
effect_size = 0.5
power = 0.8
```

Creating empty lists to fill in the effects size and power from simulated data
```{r}
effects = seq(0.1,1,0.1)
powers = seq(0.8, 0.95, 0.05)
power_alpha = array(NA, dim = c(length(5:sample), length(effects)))
new_alphas = array(NA, dim = c(length(5:sample), length(effects)))
cohensd = array(NA, dim = c(length(5:sample), length(powers)))
```

Power calculation, finding how effect size changes with power, and how power changes with effect size - for samples from 5 to 5000
#### Using different effect sizes to find power, beta and then pick between 0.05 and beta. 
```{r}
for(i in 5:sample){
  for(j in 1:length(effects)){
   
  power_alpha[i-4, j] <- pwr.t.test(power = NULL, sig.level = alpha, d = effects[j], n = i, type="paired",alternative="two.sided")$power
  beta = 1-power_alpha[i-4, j]
  new_alphas[i-4, j] = min(alpha, beta)
  }
}
```

```{r}
colnames(new_alphas) <- effects

mini_alphas <- new_alphas |> 
  as_tibble() |> 
  mutate(n = 5:sample) |> 
  pivot_longer(-n, names_to = "effect", values_to = "alpha")

colnames(power_alpha) <- effects

power_tibble_a <- power_alpha |> 
  as_tibble() |> 
  mutate(n = 5:sample) |> 
  pivot_longer(-n, names_to = "effect", values_to = "power")
```

#### Cohen's d vs new alpha levels
```{r}
for (i in 5:sample){
  for (j in 1:length(powers)){
      cohensd[i-4,j] <- pwr.t.test(power = powers[j], sig.level = alpha/i, d = NULL, n = i, type="paired",alternative="two.sided")$d
    }
  }

```

```{r}
colnames(cohensd) <- powers

mini_cohend <- cohensd |> 
  as_tibble() |> 
  mutate(n = 5:sample) |> 
  pivot_longer(-n, names_to = "power", values_to = "alpha")
```


### Graphs and Visualisations
```{r}
colours = c("#80ffea", "#8aff80", "#ffca80","#ff80bf", "#9580ff" )
bg_colours = c("#21222c", "#414558", "#a7abbe")
col_alpha = 0.6
s= 2

my_aes <- list(
  theme_dark()+
  theme(plot.background = element_rect(fill = bg_colours[1], 
                                       colour = bg_colours[1]),
        panel.background = element_rect(fill = bg_colours[1]),
        text = element_text(colour = bg_colours[3])
        ))
```

Minimizng Alpha with Beta min(alpha, beta) to form curves for 10 Cohen's D effect sizes, between 0.1 and 1. As sample size increases, as does power, meaning beta decreasing. Then once beta crosses the alpha threshold, alpha mirrors beta - meaning we achieve statistical consistency
```{r}
mini_alphas |> 
  ggplot()+
  geom_line(aes(x = n, y = alpha, colour = effect), alpha = col_alpha, size = s)+
  scale_x_log10(name = "Sample Size")+
  scale_y_continuous(name = "Alpha Level")+
  ggtitle("New Alphas vs Sample Size - For different Cohen's D")+
  my_aes
```

```{r}
mini_cohend |> 
  ggplot()+
  geom_line(aes(x = n, y = alpha, colour = power), alpha = col_alpha, size = s/2)+
  scale_x_log10(name = "Sample Size")+
  scale_y_continuous(name = "Effect Size")+
  ggtitle("New Alphas - Critical Cohen's D vs Sample Size")+
  my_aes
```
```{r}
power_tibble_a |> 
  ggplot()+
  geom_line(aes(x = n, y = power, colour = effect), alpha = col_alpha, size = s/1.5)+
  scale_x_log10(name = "Sample Size")+
  scale_y_continuous(name = "Power")+
  ggtitle("Static Alpha - Power vs Sample Size")+
  my_aes
```

A more subtle change would be to use the sample size to alter alpha, this is simpler but unclear how well it balances beta and alpha. 
```{r}
alpha_by_n <- tibble(n = 5:sample,
                     an = alpha/n,
                     an2 = alpha/n^2,
                     an3 = alpha/sqrt(n),
                     a5 = new_alphas[,5], 
                     a9 = new_alphas[,9]
                     )

alpha_by_n |> 
  ggplot()+
  geom_line(aes(x = n, y = an), colour = colours[1], alpha = col_alpha, size = s)+
  geom_line(aes(x = n, y = an2), colour = colours[2], alpha = col_alpha, size = s)+
  geom_line(aes(x = n, y = an3), colour = colours[3], alpha = col_alpha, size = s)+
  geom_line(aes(x = n, y = a5), colour = colours[4], alpha = col_alpha, size = s)+
  geom_line(aes(x = n, y = a9), colour = colours[5], alpha = col_alpha, size = s)+
  scale_x_log10(name = "Sample Size")+
  scale_y_continuous(name = "Alpha Level")+
  my_aes
```



```{r, eval = FALSE}
# powers <- seq(0.5,1,0.1)
alphas <- c(1 %o% 10^(-1:-5))
crit_d <- array(NA, dim = c(length(5:sample), length(alphas)))
power_ <- array(NA, dim = c(length(5:sample), length(alphas)))

for (i in 5:sample){
  for (j in 1:length(alphas)){
     crit_d[i-4,j] <- pwr.t.test(n = i, d = NULL, power = power, sig.level = alphas[j])$d
     power_[i-4,j] <- pwr.t.test(n = i, d = effect_size, power = NULL, sig.level = alphas[j])$power
     
  }
}
```


```{r, eval = FALSE}
colnames(crit_d) <- alphas
colnames(power_) <- alphas

eff_tibble <- crit_d |> 
  as_tibble() |> 
  mutate(n = 5:sample) |> 
  pivot_longer(cols = (-n),
               names_to = "alpha", 
               values_to = "effect")

power_tibble <- power_ |> 
  as_tibble() |> 
  mutate(n = 5:sample) |> 
  pivot_longer(cols = (-n),
               names_to = "alpha", 
               values_to = "power")

```

```{r, , eval = FALSE}

eff_tibble |> 
  ggplot(aes(x = n, y = effect, colour = as.factor(alpha)))+
  geom_line(alpha = col_alpha, size = s)+
  scale_x_log10()+
  scale_y_continuous(name = "Critical Effect Size",limits = c(0,2))+
  my_aes

```

```{r, , eval = FALSE}
power_tibble |> 
  ggplot(aes(x = n, y = power, colour = as.factor(alpha)))+
  geom_line(alpha = col_alpha, size = s)+
  scale_x_log10()+
  scale_y_continuous(name = "Power",limits = c(0,1))+
  my_aes
```



# Power curves under normality



We consider the one-sample t-test only. Results are not simulated: we use parametric predictions from the `pwr` package.



## Get power for different sample sizes and effect sizes



```{r, eval=FALSE}
nmax <- 100
nvec <- seq(10,nmax,5)
nn <- length(nvec)  
aat <- 0.05 # arbitrary alpha threshold
esvec <- seq(0, 1, 0.1)
nes <- length(esvec)
# Define array of results
power.curves <- array(NA, dim = c(nn, nes))



for(N in 1:nn){
  for(ES in 1:nes){
    power.curves[N,ES] <- pwr.t.test(n = nvec[N], d = esvec[ES], sig.level = aat)$power
  }
}
```



## Plot results



```{r}
df <- tibble(posdec = as.vector(power.curves),
             sampsize = rep(nvec, nes),
             es = factor(rep(esvec, each = nn)))



ggplot(data=df, aes(x=sampsize, y=posdec*100, colour=es)) + theme_bw() +
  geom_abline(slope = 0, intercept = 0.05*100) +
  geom_line(size=1.5) +
  geom_point(fill="white", shape=21, show.legend = FALSE) +
  scale_colour_viridis_d(option = "B", end = 0.9) +
  coord_cartesian(ylim=c(0, 100)) +
  scale_x_continuous(breaks = nvec[c(TRUE, FALSE)]) +
  scale_y_continuous(breaks = c(0, 5, seq(25,100,25))) +
  labs(x = "Sample size",
       y = "% positive decisions") +
  theme(axis.title = element_text(size = 16),
        axis.text = element_text(size = 14),
        plot.title = element_text(size = 18),
        panel.grid.minor.y = element_blank(),
        legend.title = element_text(size = 16, face = "bold"),
        legend.text = element_text(size = 12)) +
  guides(colour = guide_legend(reverse=TRUE,override.aes = list(size = 2))) +
  ggtitle('One-sample t-test power curves')



# ggsave('./figures/simres_sw_vg.png', width = 20, height = 12, units = "cm")
```
# Plot false positives as a function of sample size



With the adjustments, the corrected alpha approaches zero with increasing sample size when the null is true. With this consistent procedure, if there is no effect, none will be declared with large n.



## Compute alpha (type I error rate) for different effect sizes



We do this for the alpha = min(alpha, beta) correction.
```{r}
nmin <- 5
nmax <- 1000
nvec <- c(seq(5,10), seq(20,100,10), seq(200,1000,100))
nn <- length(nvec)  
aat <- 0.05 # arbitrary alpha threshold
esvec <- seq(0.1, 1, 0.1)
nes <- length(esvec)
sidt <- 1 # 1=1-sided, 2=2-sided
# Define array of results
fp.res <- array(NA, dim = c(nn, nes))



for(N in 1:nn){
  for(ES in 1:nes){
    power <- pwr.t.test(n = nvec[N], d = esvec[ES], sig.level = aat)$power
    beta <- 1 - power
    fp.res[N,ES] <- min(aat,beta)
  }
}
```



## Plot results
```{r}
df <- tibble(fp = as.vector(fp.res),
             sampsize = rep(nvec, nes),
             es = factor(rep(esvec, each = nn)))



ggplot(data=df, aes(x=sampsize, y=fp, colour=es)) + theme_bw() +
  geom_abline(slope = 0, intercept = 0.05*100) +
  geom_line(size=1.5) +
  geom_point(fill="white", shape=21, show.legend = FALSE) +
  scale_colour_viridis_d(option = "B", end = 0.9) +
  coord_cartesian(ylim=c(0, 0.05)) +
  # scale_x_continuous(breaks = nvec[c(TRUE, FALSE)]) +
  scale_x_continuous(trans='log10') +
  # scale_y_continuous(breaks = c(0, 5, seq(25,100,25))) +
  labs(x = "Sample size",
       y = "% positive decisions") +
  theme(axis.title = element_text(size = 16),
        axis.text = element_text(size = 14),
        plot.title = element_text(size = 18),
        panel.grid.minor.y = element_blank(),
        legend.title = element_text(size = 16, face = "bold"),
        legend.text = element_text(size = 12)) +
  guides(colour = guide_legend(reverse=TRUE,override.aes = list(size = 2))) +
  ggtitle('Alpha = min(alpha, beta)')



# ggsave('./figures/simres_sw_vg.png', width = 20, height = 12, units = "cm")
```
