---
title: "eeg-sim"
output: html_document
date: "2022-10-31"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Dependencies
```{r}
library(ggplot2)
library(tibble)
library(tidyverse)
#library(primer) # for one_over_f function
# library(permuco)
```


Edit `one_over_f` function to control variance.
Original function is available on [GitHub](https://github.com/HankStevens/primer).
Copyright Hank Stevens.

# Functions (Might Move to Seperate R scripts)
## Function to Generate 1 over F noise
```{r}
one_over_f <- function(gamma = 1, N = 200, outvar = 1){
    N.2 <- N/2
    sine.waves <- matrix(NA, nrow = N, ncol = N.2)
    steps = 2 * pi * (1:N)/N
    phase <- stats::runif(N.2, 0, 2 * pi)
    for (i in 1:N.2) {
        freq <- i
        weight <- 1/(freq^gamma)
        y <- weight * sin(freq * steps + phase[i])
        sine.waves[, i] <- y
    }
    out <- rowSums(sine.waves)
    # force var = outvar and mean = 0
    out <- out - mean(out);
    out <- out / sd(out);
    out <- out * sqrt(outvar);
    return(out)
}
```

## Function to generate raw EEG data, for two groups
```{r}

generate_raw_egg <- function(num_trials,        ## Vector or List containing size of trial(s)
                            num_time_points,    ## Number of Unique time points
                            max_time,           ## Maximum value of time, used to get freq of measurement (max/num_time_point)
                            cond1_base,         ## Template or Base used to add eeg noise to - for condition 1
                            cond2_base,         ## Template or Base used to add eeg noise to - for condition 2
                            gamma_spec_power = 1, 
                            noise_var = 1, 
                            stim_onset = 0){

  # create empty list to store df for condtions
  c_dfs <- list()

for (i in 1:length(num_trials)) {
  
  cond1 <- matrix(0, nrow = num_trials[i], ncol = num_time_points)
  cond2 <- matrix(0, nrow = num_trials[i], ncol = num_time_points)
  
  # generate trials given the number of trials
  for(t in 1:num_trials[i]){
    
    # row = trial, col = time point
    cond2[t,] <- cond2_base + one_over_f(gamma = gamma_spec_power, num_time_points, outvar = noise_var)
    cond1[t,] <- cond1_base + one_over_f(gamma = gamma_spec_power, num_time_points, outvar = noise_var)  
    
  }


  # store as tibble where cols = time, rows = trials
  c1_df <- pivot_raw_eeg(cond1, max_time, max_time/ (num_time_points - 1), stim_onset = stim_onset) |> 
    mutate(n_trial = num_trials[i], 
           cond = "cond1")
  
  c2_df <- pivot_raw_eeg(cond2, max_time, max_time/ (num_time_points - 1), stim_onset = stim_onset)|> 
    mutate(n_trial = num_trials[i],
           cond = "cond2")

  ## storing combined dfs to then bind after loop
  c_dfs[[i]] <- bind_rows(c1_df, c2_df)
  
}

## Combing tibbles to store all raw data
eeg_raw_df <- bind_rows(c_dfs)

return(eeg_raw_df)

}

```

## Function to pivot Raw Data created from generate_raw_egg()
```{r}
pivot_raw_eeg <- function(df, max_time, freq, stim_onset = 0){
    output <- df |> 
      as_tibble(.name_repair = ~ as.character(seq(0, max_time, freq))) |> 
      pivot_longer(cols = everything(), 
                   names_to = "time", 
                   values_to = "value") |> 
      mutate(time = as.integer(time),
             time = time - stim_onset)
    
    return(output)
    
  }
```

## Function to Get descriptive and inferential stats from raw EGG
```{r}
get_stats_from_raw <- function(df, alpha = 0.05){
 
  
output <- df |>  
  pivot_wider(names_from = cond, 
              values_from = value, 
              id_cols = c(time, n_trial), 
              values_fn = list) |> 
  unnest(cols = where(is.list)) |> 
  group_by(time, n_trial) |> 
  summarise(
    mn_cond1 = mean(cond1),
    mn_cond2 = mean(cond2),
    mn_diff = mn_cond2 - mn_cond1,
    p = t.test(cond1, cond2)[["p.value"]],
    ci_l = t.test(cond2, cond1, conf.level = 1 - alpha, alternative = "two.sided")[["conf.int"]][1],
    ci_u = t.test(cond2, cond1, conf.level = 1 - alpha, alternative = "two.sided")[["conf.int"]][2],
    .groups = "drop"
            )|>
  group_by(n_trial) |>
  mutate(bon_sig = if_else(p < alpha/length(unique(time)), min(ci_l), NULL),
         raw_sig = if_else(p < alpha, min(ci_l), NULL)) |>  # just so the graph has a line at zero to show significance
  ungroup()

  
return(output)
}


```

# Make template
```{r}
set.seed(314)

# true onset = 160 ms, F=17, max at F=26
true_onset <- 450  ## effect at 450ms
stim_on <- 300     ## pre-stim is 300ms
freq <- 1         ## defines resolution of stimulation, frequency of electrode
max_time <- 800
Xf <- seq(0, max_time, freq) ## start from zero and end at max time, going up in 10s (freq)
Nf <- length(Xf) 

temp1 <- vector(mode = "numeric", length = Nf)
erp <- dnorm(seq(-1.5, 1.5, length.out= 200/freq), 0, 1)
erp <- erp - min(erp)
erp <- erp / max(erp)

# getting length and storing them for later use
l_erp <- length(erp)
l_temp1 <- length(temp1)
l_pre_stim <- true_onset / freq


temp2 <- c(rep(0, l_pre_stim), erp, rep(0, (l_temp1 - l_erp - l_pre_stim)))
```

## Illustrate template
```{r}
df <- tibble(x = Xf,
             y = temp2)

ggplot(df, aes(x, y)) +
  geom_line()
```

# Generate noise

## For one gamma value
```{r}
alpha = 0.05
Nt <- 5 # number of trials
gsp <- 0.7 # gamma spectral power
resmat <- matrix(data = 0, nrow = Nt, ncol = Nf)
for(T in 1:Nt){
  resmat[T,] <- one_over_f(gamma = gsp, Nf)
}



df <- tibble(x = rep(Xf, each = Nt),
             y = as.vector(resmat),
             trials = factor(rep(seq(1,Nt), Nf)))
ggplot(df, aes(x, y, group = trials, colour = trials)) +
  geom_line()
```

## For multiple gamma values
```{r}
gsp <- seq(0, 2, 0.2) # gamma spectral power
outvar <- 1 # noise variance
Nt <- length(gsp)
resmat <- matrix(data = 0, nrow = Nt, ncol = Nf)
for(T in 1:Nt){
  set.seed(1) # same random seed to better compare trials
  resmat[T,] <- one_over_f(gamma = gsp[T], Nf, outvar = outvar)
}



df <- tibble(x = rep(Xf, each = Nt),
             y = as.vector(resmat),
             gamma = as.factor(rep(gsp, Nf)))


ggplot(df, aes(x, y, group = gamma)) +
  geom_line(size = .1) +
  theme_minimal()+
  theme(legend.position = "bottom")+
  facet_wrap(vars(gamma))
```
## Illustrate signals + noise

```{r}
# set.seed(666)
gsp <- 0.7 # gamma spectral power
outvar <- seq(0,2,0.2) # noise variance
noise_df <- list()

for(i in 1:length(outvar)){
  
  cond2 <- temp2 + one_over_f(gamma = gsp, Nf, outvar = outvar[i])
  cond1 <- temp1 + one_over_f(gamma = gsp, Nf, outvar = outvar[i])
  
  noise_df[[i]] <- tibble(x = rep(Xf,2),
             y = c(cond1, cond2),
             conds = factor(rep(c("Condition 1", "Condition 2"), each = Nf)),
             noise = outvar[i])

}

noise_df |> 
  bind_rows() |> 
  ggplot(aes(x, y, colour = conds), size = .1) +
  geom_line()+
  facet_wrap(~noise)



# Nt <- 12
# resmat.sig <- matrix(data = 0, nrow = Nt, ncol = Nf)
# resmat.noi <- matrix(data = 0, nrow = Nt, ncol = Nf)
# for(T in 1:Nt){
#   resmat.sig[T,] <- temp2 + one_over_f(gamma = gsp, Nf, outvar = outvar)
#   resmat.noi[T,] <- temp1 + one_over_f(gamma = gsp, Nf, outvar = outvar)
# }
#
# df <- tibble(x = rep(rep(Xf, each = Nt),2),
#              y = c(as.vector(resmat.sig), as.vector(resmat.noi)),
#              trials = factor(rep(1:Nt, Nf*2)),
#              cat = factor(rep(c("Signal", "Noise"), each = Nt*Nf)))
# ggplot(df, aes(x, y, group = trials, colour = cat)) +
#   geom_line(size = 1) +
#   facet_wrap(vars(trials))
```

## Generating Raw EEG Data for two cond, at various number of trials
```{r}

# set.seed(666)
Nt <- 50
Nts <- c(10,25,50,100,250,500) # number of trials
gsp <- 1 # gamma spectral power
outvar <- 1 # noise variance


eeg_raw_df <- generate_raw_egg(num_trials = Nts, 
                 num_time_points = Nf, 
                 max_time = max_time, 
                 cond1_base = temp1, 
                 cond2_base = temp2, 
                 gamma_spec_power = gsp, 
                 noise_var = outvar, 
                 stim_onset = stim_on)
```

## Illustrate averages
```{r}
eeg_df <- eeg_raw_df |>  
  get_stats_from_raw(alpha = 0.05)
```

## Plotting the Mean Differences
```{r, fig.height= 10}
## Shading is the CI of the Mean Difference

eeg_df |> 
  ggplot(aes(x = time)) +
  geom_line(aes(y = mn_cond1), colour = "#00BFC4") +
  geom_line(aes(y = mn_cond2), colour = "#F8766D") +
  geom_vline(xintercept = 0, colour = "azure4")+
  facet_wrap(~n_trial, scales =  "free")+
  theme_minimal()+
  geom_line(aes(y = bon_sig), size = 1)+
  geom_line(aes(y = raw_sig - 0.05), size = 1, colour = "deepskyblue4")+
  geom_ribbon(aes(ymax = ci_u, ymin = ci_l), alpha = .3)
```

## Plotting -log of p values, with bonferroni correction
```{r}
eeg_df |> 
  ggplot(aes(x = time, y = -log(p)))+
  # geom_point(aes(colour = p < (alpha/ Nf)), show.legend = F, size = .1)+
  geom_line() +
  geom_hline(aes(yintercept = -log(alpha/ Nf)))+
  geom_hline(aes(yintercept = -log(alpha)), colour = "azure4")+
  theme_minimal()+
  facet_wrap(~n_trial, scales = "free_y")
```

## Design two NULL ROPEs to represent control group and both groups pre-stim
## Simulation Replication of experiment with larger sample, then see if CIs are inside the rope
## Repeat with No effect between groups, and Effect between groups
## Look for false positive rate - compare raw, bonferroni p values and ROPE(s)
## Goal: Does the rope prevent FP in the null setting


## Generate Raw ERP by combining Noise and Templates
```{r}
## First study has small sample
Nt <- 25 # number of trials
gsp <- 1 # gamma spectral power
outvar <- 1 # noise variance



eeg_raw_df_trial1 <- generate_raw_egg(num_trials = Nt, 
                 num_time_points = Nf, 
                 max_time = max_time, 
                 cond1_base = temp1, 
                 cond2_base = temp2, 
                 gamma_spec_power = gsp, 
                 noise_var = outvar,
                 stim_onset = stim_on)
```

## Get stats - p-values, mean diff, quantile range and confidence intervals
```{r}
eeg_df_trial1 <- eeg_raw_df_trial1 |> 
  get_stats_from_raw(alpha = 0.05)

eeg_df_trial1 |> 
  ggplot(aes(x = time)) +
  geom_line(aes(y = mn_cond1), colour = "#00BFC4") +
  geom_line(aes(y = mn_cond2), colour = "#F8766D") +
  geom_vline(xintercept = 0, colour = "azure4")+
  theme_minimal()+
  geom_line(aes(y = bon_sig), size = 1)+
  geom_line(aes(y = raw_sig - 0.1), size = 1, colour = "deepskyblue4")+
  geom_ribbon(aes(ymax = ci_u, ymin = ci_l), alpha = .3)
```

## Design ROPEs to represent the null and alternative
```{r}
## Quantile range for pre-stim
null_quart <- quantile(x =  eeg_df_trial1 |> 
                             filter(time < 0) |> 
                             pull(mn_cond1, mn_cond2) |> 
                             as.vector(), 
                           probs = c(2*alpha, (1-2*alpha)))
    
## Quantile range for significant mean difference
alt_quart <- quantile(x = eeg_df_trial1 |> 
                        filter(p < alpha) |> 
                        pull(mn_diff) |> 
                        as.vector(), 
                      probs = c(2*alpha, (1-2*alpha)))

## Quantile range for bonferroni significant mean difference
alt_quart_bonf <- quantile(x = eeg_df_trial1 |> 
                             filter(p < (alpha/ num_time_points)) |> 
                             pull(mn_diff) |> 
                             as.vector(), 
                           probs = c(2*alpha, (1-2*alpha)))

rope <- tibble(n_trial = Nt, 
               null_min = null_quart[1], 
               null_max = null_quart[2], 
               alt_min = alt_quart[1], 
               alt_max = alt_quart[2], 
               )

eeg_df_trial1 |> 
  inner_join(rope, by = "n_trial")


```



```{r}
eeg_df_trial1 |> 
  ggplot(aes(x = time)) +
  geom_line(aes(y = mn_cond1), colour = "#00BFC4") +
  geom_line(aes(y = mn_cond2), colour = "#F8766D") +
  geom_line(aes(y = mn_diff), colour = "#000000") +
  geom_vline(xintercept = 0, colour = "azure4")+
  theme_minimal()+
  geom_line(aes(y = bon_sig), size = 1)+
  geom_line(aes(y = raw_sig - 0.1), size = 1, colour = "deepskyblue4")+
  geom_ribbon(aes(ymax = ci_u, ymin = ci_l), alpha = .3)+
  geom_ribbon(aes(x = time , ymin = rope$null_min, ymax = rope$null_max), alpha = .3, fill = "skyblue4")+
  geom_ribbon(aes(x = time , ymin = rope$alt_min, ymax = rope$alt_max), alpha = .3, fill = "darkred")
```

## Generate Raw ERP by combining Noise and Templates (Trial 2)
```{r}
## First study has small sample
Nt <- 50 # number of trials
gsp <- 1 # gamma spectral power
outvar <- 1 # noise variance

eeg_raw_df_trial2 <- generate_raw_egg(num_trials = Nt, 
                 num_time_points = Nf, 
                 max_time = max_time, 
                 cond1_base = temp1, 
                 cond2_base = temp2, 
                 gamma_spec_power = gsp, 
                 noise_var = outvar, 
                 stim_onset = stim_on)
```

## Get stats - p-values, mean diff, quantile range and confidence intervals (Trial 2)
```{r}
eeg_df_trial2 <- eeg_raw_df_trial2 |> 
  get_stats_from_raw(alpha = 0.05)

eeg_df_trial2 |> 
  ggplot(aes(x = time)) +
  geom_line(aes(y = mn_cond1), colour = "#00BFC4") +
  geom_line(aes(y = mn_cond2), colour = "#F8766D") +
  geom_line(aes(y = mn_diff), colour = "#000000") +
  geom_vline(xintercept = 0, colour = "azure4")+
  theme_minimal()+
  geom_line(aes(y = bon_sig), size = 1)+
  geom_line(aes(y = raw_sig - 0.1), size = 1, colour = "deepskyblue4")+
  geom_ribbon(aes(ymax = ci_u, ymin = ci_l), alpha = .3)+
  geom_ribbon(aes(x = time , ymin = rope$null_min, ymax = rope$null_max),alpha = .3, fill = "skyblue4")+
  geom_ribbon(aes(x = time , ymin = rope$alt_min, ymax = rope$alt_max),alpha = .3, fill = "darkred")
```
## Repeating ERPs with an effect present
```{r}
## getting pipeline function to repeat protocol
source("eeg_rope_pipeline.R")

sample_size <- 2^(1:9)

sim_df <- sim_eeg_rope_pipeline(sample_size = sample_size, 
                      alpha = 0.05, 
                      num_time_points = Nf, 
                      max_time = max_time, 
                      cond1_base = temp1, 
                      cond2_base = temp2, 
                      gamma_spec_power = gsp, 
                      noise_var = outvar, 
                      stim_onset = stim_on)
```

```{r, warning=F, fig.height=10}
sim_df |> 
  ggplot(aes(x = time)) +
  geom_line(aes(y = mn_cond1), colour = "#00BFC4", alpha = .5) +
  geom_line(aes(y = mn_cond2), colour = "#F8766D", alpha = .5) +
  geom_line(aes(y = mn_diff), colour = "#000000") +
  geom_vline(xintercept = 0, colour = "azure4")+
  geom_hline(yintercept = 0, colour = "azure4")+
  theme_minimal()+
  geom_line(aes(y = bon_sig), size = 1)+
  geom_line(aes(y = raw_sig - 0.15), size = 1, colour = "deepskyblue4")+
  geom_ribbon(aes(ymax = ci_u, ymin = ci_l), alpha = .1)+
  geom_ribbon(aes(x = time , ymin = null_min, ymax = null_max), alpha = .1, fill = "skyblue4", colour = "skyblue4")+
  # geom_ribbon(aes(x = time , ymin = alt_min, ymax = alt_max), alpha = .4, colour = "darkred", fill = "azure3")+
  geom_ribbon(aes(x = time , ymin = alt_bonf_min, ymax = alt_bonf_max), alpha = .1, fill = "darkblue",  colour = "darkblue")+
  facet_wrap(~n_trial, scale = "free_y")
```

## Repeating ERPs with no underlying effect (only noise) 
```{r}
## getting pipeline function to repeat protocol
source("eeg_rope_pipeline.R")


sim_null_df <- sim_eeg_rope_pipeline(sample_size = sample_size, 
                      alpha = 0.05, 
                      num_time_points = Nf, 
                      max_time = max_time, 
                      cond1_base = temp1, 
                      cond2_base = temp1, # both temps are the same 
                      gamma_spec_power = gsp, 
                      noise_var = outvar, 
                      stim_onset = stim_on)

```

```{r, warning=F, fig.height=10}
sim_null_df |>
  ggplot(aes(x = time)) +
  geom_line(aes(y = mn_cond1), colour = "#00BFC4") +
  geom_line(aes(y = mn_cond2), colour = "#F8766D") +
  geom_line(aes(y = mn_diff), colour = "#000000") +
  geom_vline(xintercept = 0, colour = "azure4")+
  geom_hline(yintercept = 0, colour = "azure4")+
  theme_minimal()+
  geom_line(aes(y = bon_sig), size = 1)+
  geom_line(aes(y = raw_sig - 0.15), size = 1, colour = "deepskyblue4")+
  geom_ribbon(aes(ymax = ci_u, ymin = ci_l), alpha = .1)+
  geom_ribbon(aes(x = time , ymin = null_min, ymax = null_max), alpha = .1, fill = "skyblue4", colour = "skyblue4")+
  # geom_ribbon(aes(x = time , ymin = alt_min, ymax = alt_max), alpha = .4, colour = "darkred", fill = "azure3")+
  # geom_ribbon(aes(x = time , ymin = alt_bonf_min, ymax = alt_bonf_max), alpha = .4, colour = "darkblue", fill = "azure3")+
  facet_wrap(~n_trial, scale = "free_y")
```



